---
title: "Shrinkage Methods"
author: "Ryan Yee"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```


Ridge Regression and the Lasso

**Dataset** The dataset for analysis is the Major League Baseball Data from the 1986 and 1987 seasons, which
consists of 322 observations of major league players on 20 variables including the number of hits, number of
errors, annual salary etc. Before proceeding, we first import required packages and ensure that the missing
values have been removed.

```{r}
library(ISLR)
library(glmnet)
library(dplyr)
library(tidyr)
```

Before proceeding, let's first ensure that the missing values have
been removed from the data.

```{r}
Hitters = na.omit(Hitters)
```

We will perform ridge regression and the lasso later in order to predict Salary on the Hitters data. Let’s set
up our data first.

```{r}
x = model.matrix(Salary~., Hitters)[,-1] # trim off the first column
                                         # leaving only the predictors
y = Hitters$Salary
```

The `model.matrix()` function is particularly useful for creating $x$; not only
does it produce a matrix corresponding to the 19 predictors but it also
automatically transforms any qualitative variables into dummy variables.
The latter property is important because `glmnet()` can only take numerical,
quantitative inputs.

# Ridge Regression and Lasso Regression

Ridge and lasso can be accomplished using the glmnet package. The main function in this package is glmnet(),
which can be used to fit ridge regression models, lasso models, and more. This function has slightly different
syntax from other model-fitting functions. In particular, we must pass in an x (as predictors matrix) as well
as a y (response vector), and we do not use the y ∼ x syntax.

```{r}
grid = 10^seq(10, -2, length = 100)
ridge_mod = glmnet(x, y, alpha = 0, lambda = grid)
```

By default the `glmnet()` function performs ridge regression for an automatically
selected range of $\lambda$ values. However, here we have chosen to implement
the function over a grid of values ranging from $\lambda = 10^{10}$ to $\lambda = 10^{-2}$, essentially covering the full range of scenarios from the null model containing
only the intercept, to the least squares fit. 

As we will see, we can also compute
model fits for a particular value of $\lambda$ that is not one of the original
grid values. Note that by default, the `glmnet()` function standardizes the
variables so that they are on the same scale. To turn off this default setting,
use the argument `standardize = FALSE`.

Associated with each value of $\lambda$ is a vector of ridge regression coefficients,
stored in a matrix that can be accessed by `coef()`. In this case, it is a $20 \times 100$
matrix, with 20 rows (one for each predictor, plus an intercept) and 100
columns (one for each value of $\lambda$).


We expect the coefficient estimates to be much smaller, in terms of $l_2$ norm,
when a large value of $\lambda$ is used, as compared to when a small value of $\lambda$ is
used. These are the coefficients when $\lambda = 11498$, along with their $l_2$ norm:


```{r}
ridge_mod$lambda[50] #Display 50th lambda value
coef(ridge_mod)[,50] # Display coefficients associated with 50th lambda value
sqrt(sum(coef(ridge_mod)[-1,50]^2)) # Calculate l2 norm
```

In contrast, here are the coefficients when $\lambda = 705$, along with their $l_2$
norm. Note the much larger $l_2$ norm of the coefficients associated with this
smaller value of $\lambda$.


```{r}
ridge_mod$lambda[60] #Display 60th lambda value
coef(ridge_mod)[,60] # Display coefficients associated with 60th lambda value
sum(coef(ridge_mod)[-1,60]^2) # Calculate l2 norm squared
```

We can use the `predict()` function for a number of purposes. For instance,
we can obtain the ridge regression coefficients for a new value of $\lambda$, say 50:


```{r}
pred <- pred <- predict(object = ridge_mod, newx = x)
```

Use cv.glmnet (Cross-validation) to Choose the Best Tuning Parameter

Instead of arbitrarily choosing a value of λ, it would be better to use cross-validation to choose the tuning
parameter λ. We can do this using the built-in cross-validation function, cv.glmnet(). By default, the function
performs ten-fold cross-validation, though this can be changed using the argument “folds”.

```{r}
set.seed(1) #we set a random seed first so our results will be reproducible.
cv.out.ridge=cv.glmnet(x, y, alpha = 0)
plot(cv.out.ridge)
abline(v = log(cv.out.ridge$lambda.min), col="red", lwd=3, lty=2)

```

```{r}
bestlam = cv.out.ridge$lambda.min
bestlam
```

```{r}
out = glmnet(x,y,alpha=0)
predict(out,type="coefficients",s=bestlam)[1:20,]

```

This is the best model given from the ridge regression. Notice that none of the coefficients are zero. Ridge
regression does not perform variable selection!

---

Lasso Regression
We saw that ridge regression with a wise choice of λ can outperform least squares as well as the null model on
the Hitters data set. We now ask whether the lasso can yield either a more accurate or a more interpretable
model than ridge regression. In order to fit a lasso model, we once again use the glmnet() function; however,
this time we use the argument alpha = 1. Other than that change, we proceed just as we did in fitting a
ridge model.

The lasso has the nice feature that it will set many of the coefficient estimates as exactly 0. This is useful
when some of the variables used in a multiple regression model are in fact not associated with the response.
By removing these variables (by setting the corresponding coefficient estimates to zero), we obtain a model
that is more interpretable. This is sometimes referred to as variable selection.

```{r}
grid = 10^seq(10, -2, length = 100)
lasso.mod <- glmnet(x, y, alpha=1, lambda=grid)
plot(lasso.mod, xvar="lambda", label = TRUE)
```

We can see from the coefficient plot that depending on the choice of tuning parameter, some of the coefficients
will be exactly equal to zero. We now perform cross-validation and compute the associated test error.

```{r}
set.seed(1)
cv.out.lasso = cv.glmnet(x, y, alpha = 1)
plot(cv.out.lasso)
abline(v = log(cv.out.lasso$lambda.min), col="red", lwd=3, lty=2)
```

```{r}
bestlam = cv.out.lasso$lambda.min
bestlam
```


The lasso has a substantial advantage over ridge regression in that the resulting coefficient estimates are
sparse. Here we see that some of the 19 coefficient estimates are exactly zero.

```{r}
out=glmnet(x,y,alpha=1,lambda=grid)
lasso.coef=predict(out,type="coefficients",s=bestlam)[1:20,]
lasso.coef

```